\subsection{Implementation Sections}

\begin{longtable}{|p{1cm}|p{3.5cm}|p{8.5cm}|}
\hline
\textbf{No.} & \textbf{Section Name} & \textbf{Summary (Inputs / Outputs / Properties)} \\
\hline
\endfirsthead
\hline
\textbf{No.} & \textbf{Section Name} & \textbf{Summary (Inputs / Outputs / Properties)} \\
\hline
\endhead
\hline
\endfoot
\hline
\caption{Implementation sections in the notebook}
\label{tab:implementation-sections}
\endlastfoot

1 & Import and Setup &
Inputs: none.  
Outputs: initialized environment, fixed random seed. \\
\hline

2 & Activation &
Class properties: activation function, derivative function.  
Inputs: tensor $x$.  
Outputs: activated tensor and gradient tensor. \\
\hline

3 & Loss &
Class properties: loss function, loss derivative.  
Inputs: predicted output $y_{\text{pred}}$, target $y$.  
Outputs: scalar loss value and loss gradient. \\
\hline

4 & Layers &
\texttt{FCLayer} properties: weights, bias, input data, output data.  
\texttt{ActivationLayer} properties: activation function, activation derivative.  
Inputs: input tensor.  
Outputs: transformed tensor and propagated gradient. \\
\hline

5 & Network &
Class properties: list of layers, loss function, learning rate.  
Inputs: training data and target values.  
Outputs: trained network and prediction results. \\
\hline

6 & Training Data &
Inputs: XOR input samples and corresponding labels.  
Outputs: formatted tensors for training and evaluation. \\
\hline

7 & Builder Utilities &
Inputs: layer sizes, activation types, input data.  
Outputs: constructed network instances, accuracy values, visualization figures. \\
\hline

8 & Training and Testing &
Inputs: network instance, hyperparameters.  
Outputs: accuracy metrics and decision boundary visualizations. \\
\hline

9 & Save and Load Model &
Inputs: trained network instance, file path.  
Outputs: saved model file (\texttt{xor\_model.pkl}) and loaded network instance. \\
\hline

\end{longtable}
